A word-sense representation for documents based on information value and word sense disambiguation

### 2014-03-15 paper/handpicked_table.tex
fix some spelling

\begin{tabular}{ | L{2cm} | L{5cm} |}
    \hline
    Synset & Gloss \\ \hline
    \multicolumn{2}{|c|}{Human handpicked concepts to compare against top senses} \\ \hline
            war & the waging of armed conflict against an enemy \\ \hline
            capitalism & an economic system based on private ownership of capital\\ \hline
            theory & a well-substantiated explanation of some aspect of the natural world \\ \hline
            toilet & the act of dressing and preparing yourself \\ \hline
            imperialism & a political orientation that advocates imperial interests \\ \hline
            proletarian & a member of the working class (not necessarily employed) \\ \hline
            practice & translating an idea into action \\ \hline
            revolution & a drastic and far-reaching change in ways of thinking and behaving \\ \hline
            money & the most common medium of exchange; functions as legal tender \\ \hline
            philosophy & the rational investigation of questions about existence and knowledge and ethics \\ \hline
            idealism & (philosophy) the philosophical theory that ideas are the only reality \\ \hline
            strategy & the branch of military science dealing with military command and the planning and conduct of a war \\ \hline
            materialism & (philosophy) the philosophical theory that matter is the only reality \\ \hline
            lion & large gregarious predatory feline of Africa and India having a tawny coat with a shaggy mane in the male \\ \hline
            television_receiver. & an electronic device that receives television signals and displays them on a screen \\ \hline
            hair & a covering for the body (or parts of it) consisting of a dense growth of threadlike structures (as on the human head); helps to prevent heat loss \\ \hline
            sugar & a white crystalline carbohydrate used as a sweetener and preservative \\ \hline
            car & a motor vehicle with four wheels; usually propelled by an internal combustion engine \\ \hline
\end{tabular}

>>>>

\begin{tabular}{ | L{2cm} | L{5cm} |}
    \hline
    Synset & Gloss \\ \hline
    \multicolumn{2}{|c|}{Human handpicked concepts to compare against top senses} \\ \hline
            war & the waging of armed conflict against an enemy \\ \hline
            capitalism & an economic system based on private ownership of capital\\ \hline
            theory & a well-substantiated explanation of some aspect of the natural world \\ \hline
            toilet & the act of dressing and preparing yourself \\ \hline
            imperialism & a political orientation that advocates imperial interests \\ \hline
            proletarian & a member of the working class (not necessarily employed) \\ \hline
            practice & translating an idea into action \\ \hline
            revolution & a drastic and far-reaching change in ways of thinking and behaving \\ \hline
            money & the most common medium of exchange; functions as legal tender \\ \hline
            philosophy & the rational investigation of questions about existence and knowledge and ethics \\ \hline
            idealism & (philosophy) the philosophical theory that ideas are the only reality \\ \hline
            strategy & the branch of military science dealing with military command and the planning and conduct of a war \\ \hline
            materialism & (philosophy) the philosophical theory that matter is the only reality \\ \hline
            lion & large gregarious predatory feline of Africa and India having a tawny coat with a shaggy mane in the male \\ \hline
            television\_receiver. & an electronic device that receives television signals and displays them on a screen \\ \hline
            hair & a covering for the body (or parts of it) consisting of a dense growth of threadlike structures (as on the human head); helps to prevent heat loss \\ \hline
            sugar & a white crystalline carbohydrate used as a sweetener and preservative \\ \hline
            car & a motor vehicle with four wheels; usually propelled by an internal combustion engine \\ \hline
\end{tabular}

### 2014-03-15 paper/refactor.tex
fix top senses tables
add initial results to the paper
fix some spelling

%% ADDITIONAL OPTIONAL STYLE FILES
%\usepackage[dvips]{graphicx}
%\usepackage{pnastwof}
\usepackage{amssymb,amsfonts,amsmath}
\usepackage[utf8]{inputenc}
\usepackage[activeacute, spanish]{babel}
\usepackage[dvips]{graphicx}
\usepackage{bookmark}
\usepackage{color}
\selectlanguage{spanish}

@@

But this ponderated set of top words has the problem of polysemy, tipical in the field of nlp. Polysemy occurs when a term has more than one sense or meaning, for example: the term 'bank' can refer to a financial institution or to the land formation, thus, it is polysemic. In order to resolve this limitation, we implement a basic algorithm of word-sense disambiguation based on wordnet's synsets and we apply it over the top words, using as disambiguation context the whole document. We obtain, then, the \textit{top senses} of a document, this is, a ranked set of wordnet synsets that work like a representation of the given document.

@@

\section{Results}
* Distancia a algunos conceptos de estado y revolución

\subsection{Top senses for selected works}

@@

\appendix[All tables]
\input{top_senses_the_state_and_revolution}
\appendix[Entropy of a word in a text]
\label{appendix1}
Let's recall from \cite{DARWIN} and \cite{ENTROPIC} the concept of information value of a word.
Let $T$ be a text, which has been split into $P$ pieces $T_1, T_2, \dots, T_P$, $w$ be a word in $T$,
and $f_i$ the frequency of appearance of $w$ in $T_i$.

>>>>

%% ADDITIONAL OPTIONAL STYLE FILES
%\usepackage[dvips]{graphicx}
%\usepackage{pnastwof}
\usepackage{amssymb,amsfonts,amsmath}
\usepackage[utf8]{inputenc}
\usepackage[activeacute, spanish]{babel}
\usepackage[dvips]{graphicx}
\usepackage{bookmark}
\usepackage{color}
\selectlanguage{spanish}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

@@

But this weighted set of top words has the problem of polysemy, typical in the field of nlp. Polysemy occurs when a term has more than one sense or meaning, for example: the term 'bank' can refer to a financial institution or to the land formation, thus, it is polysemic. In order to resolve this limitation, we implement a basic algorithm of word-sense disambiguation based on wordnet's synsets and we apply it over the top words, using as disambiguation context the whole document. We obtain, then, the \textit{top senses} of a document, this is, a ranked set of wordnet synsets that work like a representation of the given document.

@@

\section{Results}
\input{results}
\subsection{Top senses for selected works}

@@

\appendix[All tables]
% \input{top_senses_the_state_and_revolution}
% \input{top_senses_lenin_imperialism_the_highest_stage_of_capitalism}
% \input{top_senses_lenin_materialism_and_empirio-criticism}
% \input{top_senses_lenin_the_agrarian_programme_of_social-democracy_in_the_first_russian_revolution_1905-1907}
% \input{top_senses_lenin_what_is_to_be_done?}

\appendix[Entropy of a word in a text]
\label{appendix1}
Let's recall from \cite{DARWIN} and \cite{ENTROPIC} the concept of information value of a word.
Let $T$ be a text, which has been split into $P$ pieces $T_1, T_2, \dots, T_P$, $w$ be a word in $T$,
and $f_i$ the frequency of appearance of $w$ in $T_i$.

### 2014-03-15 paper/wsd.tex
fix some spelling

Word sense disambiguation involves the association of a given word in a text or discourse with a definition or meaning.
We found that sense disambiguation was necessary to solve in one or another level the problem of polysemy.

Our approach for WSD was a classical Lesk\cite{LESK} implementation that uses wordnet as a definition source with some tweaks.
Basically Lesk is based on the idea that words that co–occur in a sentence are being used to refer to the same topic.
The Lesk alghorithm takes the word definition or gloss that we want to disambiguate and it compares it gloss against the glosses of the other words in the context.
Then a word is assigned that sense whose gloss shares the largest number of words in common with the glosses of the other words.
This algorhitm suffers the problem that relies on the word definition and definitions usually uses as less words as possible.
Not only that sometimes two different words, but related ones, uses no words in common in their definitions.
As we use the nltk wordnet we immediately notice an improvement using the rich set of relationships that Wordnet gives to us.
We added to Lesk word stems, hypernyms, hyponyms.
For the problem of short word description we modified the algorthim to use the definition of the word to disambiguate and to compare it against the context words directly.
Finally our algorithm chooses the best synset from wordnet based on the context words.

>>>>

Word sense disambiguation involves the association of a given word in a text or discourse with a definition or meaning.

Our approach for WSD was a classical Lesk\cite{LESK} implementation that uses wordnet as a definition source with some tweaks.
Basically Lesk is based on the idea that words that co–occur in a sentence are being used to refer to the same topic.
The Lesk algorithm takes the word definition or gloss that we want to disambiguate and it compares it gloss against the glosses of the other words in the context.
Then a word is assigned that sense whose gloss shares the largest number of words in common with the glosses of the other words.
This algorithm suffers the problem that relies on the word definition and definitions usually uses as less words as possible.
Not only that sometimes two different words, but related ones, uses no words in common in their definitions.
As we use the nltk wordnet we immediately notice an improvement using the rich set of relationships that Wordnet gives to us.
We added to Lesk word stems, hypernyms, hyponyms.
For the problem of short word description we modified the algorithm to use the definition of the word to disambiguate and to compare it against the context words directly.
Finally our algorithm chooses the best synset from wordnet based on the context words.
